MSDS6371Project

### 2020-11-23 MSDS6371Project.Rmd
adding custom model

### run stepwise test and export result to csv file - Kaggle Score 0.15372
```{r}
testVariable$lSalePrice = predict(step.model,testVariable)
testVariable$SalePrice = exp(testVariable$lSalePrice)
result.stepwise = testVariable %>% dplyr::select(Id,SalePrice)
write.csv(result.stepwise,"stepwise_Prediction_Miller_YU.csv",row.names = FALSE)
```

### Forward Selection - Kaggle Score 0.15432
```{r}
set.seed(115)
#create forward fit model

#write a program to pretict 5 different combination to test forward selection model
iterations = 5
splitPerc = 0.9
total_RMSE = 0
for(i in 1:iterations){
  print(i)
  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
  train = variables.used[trainIndices,]
  test = variables.used[-trainIndices,]
  forward_fit = lm(lSalePrice~.-Id,data=train)
  forward.model = ols_step_forward_aic(forward_fit,penter=0.15)
  prediction = predict(forward.model$model,test)
  squared_MSPE = mean((test$lSalePrice - prediction)^2)
  temp_RMSE = sqrt(squared_MSPE)
  total_RMSE = total_RMSE+temp_RMSE
}
#total RMSE = 0.1235978... RMSE = 839.29 under original SalePrice scale
total_RMSE/iterations

@@

### Backward Selection - Kaggle Score 0.15432
```{r}
set.seed(112)
#create forward fit model

@@

>>>>

### run stepwise test and export result to csv file - Kaggle Score 0.15372, CV Press = 0.1494, Adjusted R2 = 0.8975
```{r}
testVariable$lSalePrice = predict(step.model,testVariable)
testVariable$SalePrice = exp(testVariable$lSalePrice)
result.stepwise = testVariable %>% dplyr::select(Id,SalePrice)
write.csv(result.stepwise,"stepwise_Prediction_Miller_YU.csv",row.names = FALSE)
```

### Forward Selection - Kaggle Score 0.15432, CV Press estimate = 0.14199, Adjusted R2 = 0.89575
```{r}
set.seed(116)
#create forward fit model

#write a program to generate 5 different combination to test forward selection model
iterations = 5
splitPerc = 0.9
total_RMSE = 0
for(i in 1:iterations){
  print(i)
  trainIndices = sample(1:dim(variables.used)[1],round(splitPerc * dim(variables.used)[1]))
  train = variables.used[trainIndices,]
  test = variables.used[-trainIndices,]
  forward_fit = lm(lSalePrice~.-Id,data=train)
  forward.model = ols_step_forward_aic(forward_fit,penter=0.15)
  prediction = predict(forward.model$model,test)
  squared_MSPE = mean((test$lSalePrice - prediction)^2)
  temp_RMSE = sqrt(squared_MSPE)
  total_RMSE = total_RMSE+temp_RMSE
}
#total RMSE = 0.1235978... 
total_RMSE/iterations

@@

### Backward Selection - Kaggle Score 0.15432, CV-Press = 0.148, adjusted R2 = 0.89575
```{r}
set.seed(112)
#create forward fit model

@@

### Custom Model - Kaggle Score = 0.14773, CV-Press = 0.1649, Adjusted R2 = 0.9207
```{r}
#reload data
amesHouse = read.csv("/Users/mingyang/Desktop/SMU/StatisticalFoundation_Fall2020/MSDS6371Project/train.csv",header = TRUE)
summary(amesHouse)
library(mice) #Load mice library to analyze the pattern of missing data
md.pattern(amesHouse)
#imputing missing data using predictive mean matching
tempData <- mice(amesHouse,m=1,maxit=50,meth='pmm',seed=20)
amesHouse = complete(tempData,1)
summary(amesHouse)

#Convert certain variables into factors
cols.to.factor = c("MSSubClass","MSZoning","Street","Alley","LotShape","LandContour","Utilities",
                   "LotConfig","LandSlope","Neighborhood","Condition1","Condition2","BldgType",
                   "HouseStyle","OverallQual","OverallCond","RoofStyle","RoofMatl","Exterior1st",
                   "Exterior2nd","MasVnrType","ExterQual","ExterCond","Foundation","BsmtQual",
                   "BsmtCond","BsmtExposure","BsmtFinType1","BsmtFinType2","Heating","HeatingQC",
                   "CentralAir","Electrical","KitchenQual","Functional","FireplaceQu","GarageType",
                   "GarageFinish","GarageQual","GarageCond","PavedDrive","PoolQC","Fence","MiscFeature",
                   "SaleType","SaleCondition","PoolArea","MoSold","YrSold","Fireplaces","FullBath")
amesHouse[cols.to.factor] = lapply(amesHouse[cols.to.factor],factor)

amesHouse$PoolYN = ifelse(is.na(amesHouse$PoolQC),"NO","YES")
amesHouse$PoolYN = as.factor(amesHouse$PoolYN)
#First test with first few variables
models1 = regsubsets(SalePrice~MSSubClass+MSZoning+LotFrontage+LotArea+Street+Alley+LotShape, data = amesHouse, nvmax = 10)
summary(models1)
#Choose the optimal model
res.sum <- summary(models1)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# As we can see among the variables selected MSSubClass, MSZoning,LotArea,Alley has variables that are significant

#Next set of variables
models2 = regsubsets(SalePrice~LandContour+Utilities+LotConfig+LandSlope+Neighborhood+Condition1+Condition2, data = amesHouse, nvmax = 15)
summary(models2)
#Choose the optimal model
res.sum <- summary(models2)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# LotConfig, Neighborhood, Condition2 has most variables that can be significant

#Next set of variables
models3 = regsubsets(SalePrice~BldgType+HouseStyle+OverallQual+OverallCond+YearBuilt, data = amesHouse, nvmax = 10)
summary(models3)
#Choose the optimal model
res.sum <- summary(models3)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# BldgType, Neighborhood, OverallQual, YearBuilt has most variables that can be significant

#Next set of variables
models4 = regsubsets(SalePrice~YearRemodAdd+RoofStyle+RoofMatl+Exterior1st+Exterior2nd, data = amesHouse, nvmax = 10)
summary(models4)
#Choose the optimal model
res.sum <- summary(models4)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# YearRemodAdd, RoofStyleHip, RoofMatl, Exterior1st, Exterior2nd has most variables that can be significant

#Next set of variables
models5 = regsubsets(SalePrice~MasVnrType+MasVnrArea+ExterQual+ExterCond+Foundation, data = amesHouse, nvmax = 10)
summary(models5)
#Choose the optimal model
res.sum <- summary(models5)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# MasVnrType, MasVnrArea, ExterQual, Foundation, have most variables that can be significant

#Next set of variables all basement section
models6 = regsubsets(SalePrice~BsmtQual+BsmtCond+BsmtExposure+BsmtFinType1+BsmtFinSF1+
                       BsmtFinType2+BsmtFinSF2+BsmtUnfSF+TotalBsmtSF, data = amesHouse, nvmax = 15)
summary(models6)
#Choose the optimal model
res.sum <- summary(models6)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# BsmtQual, BsmtFinType1, TotalBsmtSF have most variables that can be significant

#Next set of variables  - Utilities
models7 = regsubsets(SalePrice~Heating+HeatingQC+CentralAir+Electrical, data = amesHouse, nvmax = 10)
summary(models7)
#Choose the optimal model
res.sum <- summary(models7)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# Heating, HeatingQC, CentralAir,Elecrical have most variables that can be significant

#Next set of variables  - Square foots
models8 = regsubsets(SalePrice~X1stFlrSF+X2ndFlrSF+LowQualFinSF+GrLivArea+BsmtFullBath+BsmtHalfBath+FullBath+HalfBath, data = amesHouse, nvmax = 10)
summary(models8)
#Choose the optimal model
res.sum <- summary(models8)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# X1stFlrSF, X2ndFlrSF, LowQualFinSF,GrLivArea,BsmtFullBath,BsmtHalfBath,FullBath1,HalfBath have most variables that can be significant

#Next set of variables  - Square foots
models9 = regsubsets(SalePrice~BedroomAbvGr+KitchenAbvGr+KitchenQual+TotRmsAbvGrd+Functional, data = amesHouse, nvmax = 10)
summary(models9)
#Choose the optimal model
res.sum <- summary(models9)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# BedroomAbvGr, KitchenAbvGr, KitchenQual,TotRmsAbvGrd have most variables that can be significant

#Next set of variables  
models10 = regsubsets(SalePrice~Fireplaces+FireplaceQu+GarageType+GarageYrBlt+
                        GarageFinish+GarageCars+GarageArea+GarageQual+GarageCond, data = amesHouse, nvmax = 10)
summary(models10)
#Choose the optimal model
res.sum <- summary(models10)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# Fireplaces, FireplaceQu, GarageFinish,GarageCars,GarageArea have most variables that can be significant

#Next set of variables  
models11 = regsubsets(SalePrice~PavedDrive+WoodDeckSF+OpenPorchSF+EnclosedPorch+
                        X3SsnPorch+ScreenPorch, data = amesHouse, nvmax = 10)
summary(models11)
#Choose the optimal model
res.sum <- summary(models11)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# PavedDriveY, WoodDeckSF, OpenPorchSF,X3SsnPorch,ScreenPorch have most variables that can be significant

#Next set of variables  
models12 = regsubsets(SalePrice~PoolYN+Fence+MiscVal+MoSold+YrSold, data = amesHouse, nvmax = 10)
summary(models12)
#Choose the optimal model
res.sum <- summary(models12)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)
# PoolYN, Fence, MoSold,YrSold have most variables that can be significant

#Select Variables
custom.var = amesHouse %>% 
  dplyr::select(Id,MSSubClass, MSZoning,LotArea,LotConfig, Neighborhood, Condition2,
                BldgType, Neighborhood, OverallQual, YearBuilt,YearRemodAdd, RoofStyle, RoofMatl, Exterior1st, Exterior2nd,
                MasVnrType, MasVnrArea, ExterQual, Foundation, BsmtQual, BsmtFinType1, TotalBsmtSF,
                Heating, HeatingQC, CentralAir,Electrical,
                X1stFlrSF, X2ndFlrSF, LowQualFinSF,GrLivArea,BsmtFullBath,BsmtHalfBath,FullBath,HalfBath,
                BedroomAbvGr, KitchenAbvGr, KitchenQual,TotRmsAbvGrd,
                Fireplaces,GarageCars,GarageArea,
                PavedDrive, WoodDeckSF, OpenPorchSF,X3SsnPorch,ScreenPorch,
                PoolYN, MoSold,YrSold,
                SalePrice)
custom.var$SalePrice = log(custom.var$SalePrice)
custom.var$GrLivArea = log(custom.var$GrLivArea)

#Adjusting NA factors
custom.var$MasVnrType = as.character(custom.var$MasVnrType)
custom.var$MasVnrType = ifelse(is.na(custom.var$MasVnrType),"NONE",custom.var$MasVnrType)
custom.var$MasVnrType = as.factor(custom.var$MasVnrType)

custom.var$BsmtQual = as.character(custom.var$BsmtQual)
custom.var$BsmtQual = ifelse(is.na(custom.var$BsmtQual),"NONE",custom.var$BsmtQual)
custom.var$BsmtQual = as.factor(custom.var$BsmtQual)

custom.var$BsmtFinType1 = as.character(custom.var$BsmtFinType1)
custom.var$BsmtFinType1 = ifelse(is.na(custom.var$BsmtFinType1),"Unf",custom.var$BsmtFinType1)
custom.var$BsmtFinType1 = as.factor(custom.var$BsmtFinType1)

custom.var$Electrical = as.character(custom.var$Electrical)
custom.var$Electrical = ifelse(is.na(custom.var$Electrical),"SBrkr",custom.var$Electrical)
custom.var$Electrical = as.factor(custom.var$Electrical)

#Model Training
set.seed(11)
train.control <- trainControl(method = "cv", number = 10)
custom.model = train(SalePrice~.-Id, data=custom.var,
                   method="lmStepAIC",
                   trControl = train.control,
                   trace=FALSE)
custom.model$results
custom.model$finalModel
summary(custom.model$finalModel)
print(custom.model)
# RMSE = 0.1494448 by 10 fold internal cross-validation

ols_plot_resid_stand(custom.model$finalModel)
ols_plot_cooksd_bar(custom.model$finalModel)
plot(custom.model$finalModel)
#Load in testing data and fill in the missing data in the same manner
testData = read.csv("/Users/mingyang/Desktop/SMU/StatisticalFoundation_Fall2020/MSDS6371Project/test.csv",header = TRUE)
testData$MSSubClass = ifelse(testData$MSSubClass==150,180,testData$MSSubClass)
testData$FullBath = ifelse(testData$FullBath==4,3,testData$FullBath)
testData$Fireplaces = ifelse(testData$Fireplaces==4,3,testData$Fireplaces)
testData$MasVnrArea = ifelse(is.na(testData$MasVnrArea),0,testData$MasVnrArea)
testData$TotalBsmtSF = ifelse(is.na(testData$TotalBsmtSF),0,testData$TotalBsmtSF)
testData$BsmtFullBath = ifelse(is.na(testData$BsmtFullBath),0,testData$BsmtFullBath)
testData$BsmtHalfBath = ifelse(is.na(testData$BsmtHalfBath),0,testData$BsmtHalfBath)
testData$GarageCars = ifelse(is.na(testData$GarageCars),0,testData$GarageCars)
testData$GarageArea = ifelse(is.na(testData$GarageArea),0,testData$GarageArea)
testData[cols.to.factor] = lapply(testData[cols.to.factor],factor)

testData$PoolYN = ifelse(is.na(testData$PoolQC),"NO","YES")
testData$PoolYN = as.factor(testData$PoolYN)
summary(testData)
#Testing test set
testData1 = testData %>% 
  dplyr::select(Id,MSSubClass, MSZoning,LotArea,LotConfig, Neighborhood, Condition2,
                BldgType, Neighborhood, OverallQual, YearBuilt,YearRemodAdd, RoofStyle, RoofMatl, Exterior1st, Exterior2nd,
                MasVnrType, MasVnrArea, ExterQual, Foundation, BsmtQual, BsmtFinType1, TotalBsmtSF,
                Heating, HeatingQC, CentralAir,Electrical,
                X1stFlrSF, X2ndFlrSF, LowQualFinSF,GrLivArea,BsmtFullBath,BsmtHalfBath,FullBath,HalfBath,
                BedroomAbvGr, KitchenAbvGr, KitchenQual,TotRmsAbvGrd,
                Fireplaces,GarageCars,GarageArea,
                PavedDrive, WoodDeckSF, OpenPorchSF,X3SsnPorch,ScreenPorch,
                PoolYN, MoSold,YrSold)
testData1$GrLivArea = log(testData1$GrLivArea)
summary(testData1)
#Adjusting NA factors
testData1$MSZoning = as.character(testData1$MSZoning)
testData1$MSZoning = ifelse(is.na(testData1$MSZoning),"RL",testData1$MSZoning)
testData1$MSZoning = as.factor(testData1$MSZoning)

testData1$Exterior1st = as.character(testData1$Exterior1st)
testData1$Exterior1st = ifelse(is.na(testData1$Exterior1st),"Other",testData1$Exterior1st)
testData1$Exterior1st = as.factor(testData1$Exterior1st)

testData1$Exterior2nd = as.character(testData1$Exterior2nd)
testData1$Exterior2nd = ifelse(is.na(testData1$Exterior2nd),"Other",testData1$Exterior2nd)
testData1$Exterior2nd = as.factor(testData1$Exterior2nd)

testData1$BsmtQual = as.character(testData1$BsmtQual)
testData1$BsmtQual = ifelse(is.na(testData1$BsmtQual),"TA",testData1$BsmtQual)
testData1$BsmtQual = as.factor(testData1$BsmtQual)

testData1$BsmtFinType1 = as.character(testData1$BsmtFinType1)
testData1$BsmtFinType1 = ifelse(is.na(testData1$BsmtFinType1),"Unf",testData1$BsmtFinType1)
testData1$BsmtFinType1 = as.factor(testData1$BsmtFinType1)

testData1$KitchenQual = as.character(testData1$KitchenQual)
testData1$KitchenQual = ifelse(is.na(testData1$KitchenQual),"TA",testData1$KitchenQual)
testData1$KitchenQual = as.factor(testData1$KitchenQual)

testData1$MasVnrType = as.character(testData1$MasVnrType)
testData1$MasVnrType = ifelse(is.na(testData1$MasVnrType),"None",testData1$MasVnrType)
testData1$MasVnrType = as.factor(testData1$MasVnrType)

testData1$Exterior1st = as.character(testData1$Exterior1st)
testData1$Exterior1st = ifelse(testData1$Exterior1st=="Other","VinylSd",testData1$Exterior1st)
testData1$Exterior1st = as.factor(testData1$Exterior1st)

testData1$SalePrice = predict(custom.model,testData1)
testData1$SalePrice = exp(testData1$SalePrice)
custom.result = testData1 %>% dplyr::select(Id,SalePrice)
write.csv(custom.result,"custom_model_Miller_YU.csv",row.names = FALSE)
```
